Promotors' reports on the first draft of the thesis document.

### 2020-08-21 ChapterRampInput/ChapterRampInput.tex
Final final version before internal defense
Second draft before public defense

The goal of any measurement is accurately estimating the input using the response of the sensor in the transient regime.
Here we consider the case where the input varies during the measurement.
The simplest case is when the input changes with a constant rate, such as being described with an affine function.
The inputs that vary at a constant rate activate the sensor gradually. 
An affine input estimation method is needed to estimate the slope and the intercept parameters of the affine function.

This chapter describes two signal processing methods for estimating affine input parameters.
These methods are a data-driven subspace method and a maximum-likelihood (ML) estimation method based on local optimization.
The subspace is a recursive method independent of the sensor model, and the ML method iteratively minimizes a cost function using a simulated sensor response in a receding time horizon.

In this chapter, the dynamic weighing is an implementation example.
The dynamic weighing of an affine input is a demanding metrology application because the affine excitation turns the linear time-invariant weighing sensor into a time-varying system.
Then, the affine input estimation methods provide insight into the properties of the dynamic measurement of time-varying inputs.

The effectiveness and uncertainty of the proposed methods are evaluated with simulation studies.
The uncertainty is assessed, for the adaptive data-driven method, using a Taylor series expansion of the exponentially weighted recursive least-squares estimate, and
for the ML method, using the derivatives of the residual error obtained from the to-be-minimized cost function.

@@

\paragraph{Problem} 
Given the sequence of measured output observations $\widetilde{\mathbf{y}} = \left( \widetilde{y}(1),\ldots, \widetilde{y}(T)\right)$, with $\widetilde{y}(t) \in {\rm I\!R}$, of a stable linear time-invariant system of order $n$, and static gain $\mathbf{G}$, generated by an affine input $u(t) = \widebar{a} t + \widebar{b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u}(t) = \widehat{a} t + \widehat{b}$ approximates $u$(t).
The measured observations $\widetilde{\mathbf{y}} = \widebar{\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses $\widebar{\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.

@@

\node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
\node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u(t)$};

@@

The application of an affine input turns the linear time-invariant system into a linear time-varying system, whose
dynamics depends on the input $u(t) = \widebar{a} t + \widebar{b}$, as it is described by the differential equation:
\begin{equation} \dfrac{d}{dt} \left( \left( \widebar{a} t + \widebar{b} + m \right) \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left( \widebar{a} t + \widebar{b} + m \right) g \end{equation}
where $m$ is the mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.

The weighing system admits a state-space representation where the states $x_1=y$ and $x_2=\dot{y}$ are the position and the speed of the weighing scale: 
\begin{equation} \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{\widebar{a} t + \widebar{b} + m} & \dfrac{-(k_{\mathrm{d}} + \widebar{a})}{\widebar{a} t + \widebar{b} + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix}, \quad y = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} . \label{eqn:ltvsystem} \end{equation}

@@

\subsection{Subspace method}
% \linebreak
The errors-in-variables (EIV) minimization problem (\ref{eqn:min_seiv}) is extended to estimate the affine input parameters.
The profile of an affine input excitation can be reconstructed by solving the structured EIV problem  
\begin{equation} \widehat{\bm{\theta}} = \underset{\bm{\theta}}{\mathrm{argmin}} \ \left\Vert \bm{\Lambda}^{1/2} \left( \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta} \right) \right\Vert^2_2 . \label{eqn:min_ewrls} \end{equation}
where $\widetilde{\mathbf{y}}$ is the observed affine response, $\widetilde{\mathbf{K}}$ is defined in (\ref{eqn:K0plusnoise}), and
the matrix $\bm{\Lambda} \in \mathbb{R}^{(T-n) \times (T-n)}$ is a diagonal matrix of descending powers of the weight factor $\lambda \in [0, 1)$, \textit{i.e.},  $\bm{\Lambda} = \mathrm{diag}\left(\lambda^{T-n}, \ldots, \lambda^2, \lambda^1 \right)$.
The weight factor $\lambda$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta}$.

The exponentially weighted RLS solution to the problem (\ref{eqn:min_ewrls}) is implemented to facilitate the online implementation of the affine input estimation method.
When $\lambda=1$, the exponentially weighted RLS solution is identical to that of the RLS.
When $\lambda<1$, the older residuals are weighted with lower values than the residuals of recent observations.
In this way, the solution of the minimization problem depends more on newer data, and less on older data.

In Figure \ref{fig:sensor_response} we can see an example of the affine input $\mathbf{u}$ that excites a weighing sensor, and the generated response $\widetilde{\mathbf{y}}$.
After the transient effects have diminished, the sensor response is a ramp of the same slope as the input, but with a different intercept due to an offset $\tau$.
The offset $\tau$ is explained by the parameter $m$ in the sensor representation given in Figure \ref{fig:msd_system}.
This is the mass of the sensor itself, that is added to the mass of the object under measurement.	
From the point of view of the subspace method, the value of $m$ is unknown, and there is no interest in its estimation.

The exponentially weighted RLS solution to the problem (\ref{eqn:min_ewrls}) provides an affine input  estimation $\widehat{\mathbf{u}}$. 
The affine input parameters $a$ and $b$ are estimated, in a second step, by fitting $\widehat{\mathbf{u}}$ to a straight line using linear regression, as follows
\begin{equation} 
  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ T & 1 \end{bmatrix} 
  \begin{bmatrix} \widehat{a} \\ \widehat{b}  \end{bmatrix} = 
  \begin{bmatrix} \widehat{u}(1)-\tau \\ \vdots \\ \widehat{u}(T)-\tau \end{bmatrix}, \label{eqn:LS}
\end{equation}
where it is considered the existance of the offset $\tau$.

For each new observation $y(t)$, the estimation $\widehat{\mathbf{u}}$ is updated followed by the RLS update of the slope $\widehat{a}$ and the intercept $\widehat{b}$.
The values of the tuning parameters $\lambda$ and $\tau$ can be obtained in the calibration of the method using the response of the sensor.

The calibration procedure consists in the measurement of an object with a reference mass.
During the calibration, $\lambda$ and $\tau$ are tuned by changing their values aiming to reduce the estimation errors of $\widehat{a}$ and $\widehat{b}$.
This reduction of the estimation errors is observed in the convergence of the affine input parameters towards their true values.

\begin{figure}[!htbp]
\centering
\includegraphics[width=1\columnwidth]{./ChapterRampInput/fig/Fig_2_1.pdf} 
\caption{ \label{fig:sensor_response} When the transient part of the sensor response $\widetilde{\mathbf{y}}$ asymptotically decreases, a ramp with the same slope $a$ of the  affine input $u(t) = at+b$ predominates. The intercept of the prevailing ramp is the intercept $b$ added to and offset $\tau$.}
\end{figure}

@@

To obtain the first and second moments of the step input estimate $\widehat{\mathbf{u}}$, we need to study the solution
\begin{equation} \widehat{\bm{\theta}} = ( \widetilde{\mathbf{K}}^\top \bm{\Lambda} \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \bm{\Lambda} \widetilde{\mathbf{y}} , \label{eqn:xhatW} \end{equation}
of the overdetermined structured errors-in-variables EIV problem (\ref{eqn:min_ewrls}).
Using a second order Taylor series expansion of the inverse matrix we can approximate the LS solution as
\begin{equation} \widehat{\bm{\theta}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\mathbf{K} + \mathbf{E})^\top \bm{\Lambda} (\mathbf{y} + \bm{\epsilon}). \label{eqn:xhatexpW} \end{equation} 
where 
\begin{equation} \mathbf{C} = \mathbf{K}^\top \bm{\Lambda} \mathbf{K}, \quad \text{and} \quad \mathbf{M} = \mathbf{C}^{-1} ( \mathbf{K}^\top \bm{\Lambda} \mathbf{E} + \mathbf{E}^\top \bm{\Lambda} \mathbf{K} + \mathbf{E}^\top \bm{\Lambda} \mathbf{E} ). \end{equation}

The Taylor series approximation of $\widehat{\bm{\theta}}$ enables the calculation of the estimation bias and covariance since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion. 
The bias and the covariance of the estimate $\widehat{\bm{\theta}}$ are obtained from the definitions (\ref{eqn:biasdef}), and (\ref{eqn:covdef}). %The bias and the covariance of the estimate $\widehat{\bm{\theta}}$ are obtained from \begin{equation}  \mathbf{b} \left(\widehat{\bm{\theta}} \right) = \mathbf{\mu} - \widebar{\bm{\theta}}, \label{eqn:biasdef} \end{equation}  \begin{equation} \begin{aligned} \mathrm{\mathbf{Cov}} \left( \widehat{\bm{\theta}} \right) & = \mathbb{E} \left\{ \left( \widehat{\bm{\theta}} - \mathbf{\mu} \right)  \left( \widehat{\bm{\theta}} - \mathbf{\mu} \right)^\top \right\} . \end{aligned} \label{eqn:covdef} \end{equation} where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\}$ is the expected value of the estimate and $\widebar{\bm{\theta}}$ is the true value. 
Considering the structure of the EIV problem, the bias and the covariance of the estimate approximation (\ref{eqn:xhatexpW}) can be expressed as
\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \bm{\Lambda} \mathbf{B}_1 - \mathbf{B}_2 \right) \bm{\theta} - \left( \mathbf{K}^\top \bm{\Lambda} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasEW} \end{equation}
\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \bm{\Lambda} \left( \sigma_{\epsilon}^2 \mathbf{I}_{T-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \bm{\Lambda} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right), \end{aligned} \label{eqn:varEW} \end{equation}
where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\Lambda} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Lambda} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\Lambda} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Lambda} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \bm{\Lambda}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.

>>>>

\color{blue} 
In this thesis document we have discussed that a measurement estimates the unknown value of a physical quantity, also called measurand.
The physical quantity of interest is applied as an input signal to a dynamic system, the sensor, and its state changes from the initial conditions into a transient state.
The input value is estimated using the sensor response in transient state.
A dynamic measurement occurs when the input fluctuations impacts on the input value estimation.
A typical example of a dynamic measurement problem is a low-bandwidth sensor excited with a fast changing input as is explained in \citet{Dienstfrey14}.
In this case, the sensor transient response is slower than the variation of the input, making it difficult to track the input evolution.
The variation of some input characteristics, like the minimum or maximum or the effects of the environment, are considered in applications such as measurements of temperature in \citet{Saggin01}, pressure in \citet{Matthews14}, acceleration in \citet{Link07}, force in \citet{Vlajic16, Hessling08a}, and mass in \cite{Shu93, Boschetti13}.

The data-driven step input estimation method, introduced by \citet{Markovsky15cep} and described in Chapter \ref{chap:Preliminaries},
estimates the unknown step level of an input directly from the sensor step response.
Since most inputs have variations, contrary to the step response that is constant after its application, sets the need for methods for estimating more input models.
One of such input models are affine inputs that vary at a constant rate, and they are found in applications where the measurand activates the sensor gradually. 
An example of gradual activation is the mass measurement of objects in motion, that exists in along conveyor belts.

Current solutions to the weighing in motion are low pass filters that estimate the mass using a saturated ramp as it is proposed by \citet{Tasaki07}, and by \citet{Niedzwiecki16b}, and also time varying filters are developed in \citet{Piskorowski08} and in \citet{Pietrzak14}.
Nevertheless, the signal processing affine input estimation methods are motivated by the need to obtain the mass of the object from the ramp before it reaches saturation.
The ramp is parameterized as a straight line model where the slope and the interception are the parameters of interest of an affine input model.
This chapter describes two methods for the estimation of the affine input parameters.

One method is an extension of the data-driven step input estimation method, and then it is a recursive algorithm of low computational cost that can be implemented in real-time.
The second method is a maximum-likelihood (ML) estimation method based on local-optimization that is proposed for off-line processing of the sensor transient response.
The ML method is similar to the model predictive control approach, as discussed in \cite{Mayne14}, because a cost function is minimized iteratively to optimize the parameters of a sensor model using the observed sensor response in a receding time horizon.
The difference is that the ML method aims to estimate the unknown value of the affine input parameters instead of identifying a model and controlling the dynamic system.
Nevertheless, the ML method can estimate the parameters of the affine input, the parameters of a sensor model, and the initial conditions of the sensor.

The uncertainty of the subspace method is assessed in \cite{QuintanaCSDA} using a Taylor expansion of the estimate and Monte Carlo random sampling approach.
The Monte Carlo approach requires a large set of generated random samples, and according to \citet{Cox06}, for simple systems it is the recommended method.
The uncertainty of the ML method is assessed using the derivatives of the residual error that constructs the to-be-minimized cost function.
The covariance of the optimization method estimate is found in \cite{Pintelon12Book} using the inverse of the Hessian matrix.
\color{black}

@@

\paragraph{Problem} 
Given the sequence of measured output observations $\widetilde{\mathbf{y}} = \left( \widetilde{y}(1),\ldots, \widetilde{y}(\color{blue}N\color{black})\right)$, with $\widetilde{y}(t) \in \mathbb{R}$, of a stable linear time-invariant system of order $n$, and static gain ${G}$, generated by an affine input $u(t) = {a} t + {b}$, estimate the parameters of the affine input, \textit{i.e.,} find the values of the parameters $\widehat{a}, \widehat{b} \in {\rm I\!R}$ such that $\widehat{u}(t) = \widehat{a} t + \widehat{b}$ approximates $u$(t).
The measured observations $\widetilde{\mathbf{y}} = {\mathbf{y}} + \bm{\epsilon}$ are exact sensor responses ${\mathbf{y}}$ perturbed by additive noise  $\bm{\epsilon}$ assumed to be independent and normally distributed of zero mean and given variance $\sigma_{\epsilon}^2$.

@@

\node (M) [minimum width=2.5cm,minimum height=0.05cm] {$m$};
\node (Mu) [minimum width=2.5cm,minimum height=0.75cm,yshift=0.57cm] {$u=at+b$};

@@

The application of an affine input turns the linear time-invariant system into a linear time-varying system, whose
dynamics depends on the input $u(t) = {a} t + {b}$, as it is described by the differential equation:
\begin{equation} \dfrac{d}{dt} \left( \left( {a} t + {b} + m \right) \dfrac{dy}{dt} \right) + k_{\mathrm{d}} \dfrac{dy}{dt} + k_{\mathrm{s}} y = \left( {a} t + {b} + m \right) g \end{equation}
where $m$ is the mass of the scale, $k_{\mathrm{d}}$ is the damping constant, $k_{\mathrm{s}}$ is the elasticity constant, and $g = 9.81$ $\mathrm{m/s}^2$ is the gravitational acceleration.

The weighing system admits a state-space representation where the states $x_1=y$ and $x_2=\dot{y}$ are the position and the speed of the weighing scale: 
\begin{equation} \dot{\mathbf{x}} = \begin{bmatrix} 0 & 1 \\ \dfrac{-k_{\mathrm{s}}}{{a} t + {b} + m} & \dfrac{-(k_{{d}} + {a})}{{a} t + {b} + m} \end{bmatrix} \mathbf{x} + \begin{bmatrix} 0 \\ g \end{bmatrix}, \quad y = \begin{bmatrix} 1 & 0  \end{bmatrix} \mathbf{x} . \label{eqn:ltvsystem} \end{equation}

@@

\subsection{Subspace method}
% \linebreak
The errors-in-variables (EIV) minimization problem (\ref{eqn:min_seiv}) is extended to estimate the affine input parameters.
The profile of an affine input excitation can be reconstructed by solving the structured EIV problem  
\begin{equation} \widehat{\bm{\theta}} = \underset{\bm{\theta}}{\mathrm{argmin}} \ \left\Vert \bm{\Omega}^{1/2} \left( \widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta} \right) \right\Vert^2_2 . \label{eqn:min_ewrls} \end{equation}
where $\widetilde{\mathbf{y}}$ is the observed affine response, $\widetilde{\mathbf{K}}$ is defined in (\ref{eqn:K0plusnoise}), and
the matrix $\bm{\Omega} \in \mathbb{R}^{(\color{blue}N\color{black}-n) \times (\color{blue}N\color{black}-n)}$ is a diagonal matrix of descending powers of the weight factor $\omega \in [0, 1)$, \textit{i.e.},  $\bm{\Omega} = \mathrm{diag}\left(\omega^{\color{blue}N\color{black}-n}, \ldots, \omega^2, \omega^1 \right)$.
The weight factor $\omega$ is a data selection forgetting factor since it enables to apply different weights to the residuals $\widetilde{\mathbf{y}} - \widetilde{\mathbf{K}} \bm{\theta}$.

The exponentially weighted RLS solution to the problem (\ref{eqn:min_ewrls}) is implemented to facilitate the online implementation of the affine input estimation method.
\color{blue}
The exponentially weighted recursive least-squares (RLS) can solve recursively the estimation problem (\ref{eqn:min_seiv}).
In \citet{Kailath00book} the exponentially weighted RLS algorithm is described as follows

\begin{equation} \begin{aligned} \widehat{\bm{\theta}}(k) &= \widehat{\bm{\theta}}(k-1) + \bm{\kappa}_{k} \left( \widetilde{y}(k) - \widetilde{\mathbf{k}}_{k} \widehat{\bm{\theta}}(k-1) \right) , \\  \bm{\kappa}_{k} &= \omega^{-1} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top / \left( 1 + \omega^{-1} \widetilde{\mathbf{k}}_{k} \bm{\Psi}(k-1) \widetilde{\mathbf{k}}_{k}^\top  \right) \\ \bm{\Psi}(k) &= \omega^{-1} \left( \mathbf{I} - \bm{\kappa}_{k} \widetilde{\mathbf{k}}_{k} \right) \bm{\Psi}(k-1), \label{eqn:RLS} \end{aligned} \end{equation}
for $k = 2n+1, 2n+2, \ldots$, where $\widetilde{\mathbf{k}}_{k}$ represents the row of $\widetilde{\mathbf{K}}$ that corresponds to the $k\mathrm{-th}$ sample, $\bm{\kappa}_{k}$ is a gain scalar, and $\Psi(k)$ is a covariance matrix.
The estimates and the covariance matrix are initialized using the first $n+1$ samples, i.e.,  
\begin{equation} \begin{aligned} \widehat{\bm{\theta}} (2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top  \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega}  \widetilde{\mathbf{y}}_{n+1}, \quad \mathrm{and} \\ \Psi(2n+1) &= \left( \widetilde{\mathbf{K}}_{n+1}^\top \bm{\Omega} \widetilde{\mathbf{K}}_{n+1} \right)^{-1} , \end{aligned} \end{equation} 
where $\widetilde{\mathbf{K}}_{n+1}$ is the matrix $\widetilde{\mathbf{K}}$ with the first $n+1$ rows, and $\widetilde{\mathbf{y}}_{n+1}$ is the vector $\widetilde{\mathbf{y}}$ with the first $n+1$ elements\color{black}.
When $\omega=1$, the exponentially weighted RLS solution is identical to that of the RLS.
When $\omega<1$, the older residuals are weighted with lower values than the residuals of recent observations.
In this way, the solution of the minimization problem depends more on newer data, and less on older data.

In Figure \ref{fig:sensor_response} we can see an example of the affine input $\mathbf{u}$ that excites a weighing sensor, and the generated response $\widetilde{\mathbf{y}}$.
After the transient effects have diminished, the sensor response is a ramp of the same slope as the input, but with a different intercept due to an offset $\omicron$.
The offset $\omicron$ is explained by the parameter $m$ in the sensor representation given in Figure \ref{fig:msd_system}.
This is the mass of the sensor itself, that is added to the mass of the object under measurement.	
From the point of view of the subspace method, the value of $m$ is unknown, and there is no interest in its estimation.

The exponentially weighted RLS solution to the problem (\ref{eqn:min_ewrls}) provides an affine input  estimation $\widehat{\mathbf{u}}$. 
The affine input parameters $a$ and $b$ are estimated, in a second step, by fitting $\widehat{\mathbf{u}}$ to a straight line using linear regression, as follows
\begin{equation} 
  \begin{bmatrix} 1 & 1  \\ \vdots & \vdots \\ \color{blue}N\color{black} & 1 \end{bmatrix} 
  \begin{bmatrix} \widehat{a} \\ \widehat{b}  \end{bmatrix} = 
  \begin{bmatrix} \widehat{u}(1)-\omicron \\ \vdots \\ \widehat{u}(\color{blue}N\color{black})-\omicron \end{bmatrix}, \label{eqn:LS}
\end{equation}
where it is considered the existance of the offset $\omicron$.

For each new observation $y(t)$, the estimation $\widehat{\mathbf{u}}$ is updated followed by the RLS update of the slope $\widehat{a}$ and the intercept $\widehat{b}$.
The values of the tuning parameters $\omega$ and $\omicron$ can be obtained in the calibration of the method using the response of the sensor.

The calibration procedure consists in the measurement of an object with a reference mass.
During the calibration, $\omega$ and $\omicron$ are tuned by changing their values aiming to reduce the estimation errors of $\widehat{a}$ and $\widehat{b}$.
This reduction of the estimation errors is observed in the convergence of the affine input parameters towards their true values.

\begin{figure}[!htbp]
\centering
\includegraphics[width=0.69\columnwidth]{./ChapterRampInput/fig/Fig_2_1.pdf} 
\caption{ \label{fig:sensor_response} When the transient part of the sensor response $\widetilde{\mathbf{y}}$ asymptotically decreases, a ramp with the same slope $a$ of the  affine input $u(t) = at+b$ predominates. The intercept of the prevailing ramp is the intercept $b$ added to and offset $\omicron$.}
\end{figure}

@@

To obtain the first and second moments of the step input estimate $\widehat{\mathbf{u}}$, we need to study the solution
\begin{equation} \widehat{\bm{\theta}} = ( \widetilde{\mathbf{K}}^\top \bm{\Omega} \widetilde{\mathbf{K}}  )^{-1} \widetilde{\mathbf{K}}^\top \bm{\Omega} \widetilde{\mathbf{y}} , \label{eqn:xhatW} \end{equation}
of the overdetermined structured errors-in-variables EIV problem (\ref{eqn:min_ewrls}).
Using a second order Taylor series expansion of the inverse matrix we can approximate the LS solution as
\begin{equation} \widehat{\bm{\theta}} \approx \left( \mathbf{I} - \mathbf{M} + \mathbf{M}^2 \right) \mathbf{C}^{-1} (\mathbf{K} + \mathbf{E})^\top \bm{\Omega} (\mathbf{y} + \bm{\epsilon}). \label{eqn:xhatexpW} \end{equation} 
where 
\begin{equation} \mathbf{C} = \mathbf{K}^\top \bm{\Omega} \mathbf{K}, \quad \text{and} \quad \mathbf{M} = \mathbf{C}^{-1} ( \mathbf{K}^\top \bm{\Omega} \mathbf{E} + \mathbf{E}^\top \bm{\Omega} \mathbf{K} + \mathbf{E}^\top \bm{\Omega} \mathbf{E} ). \end{equation}

The Taylor series approximation of $\widehat{\bm{\theta}}$ enables the calculation of the estimation bias and covariance since the measurement noise $\bm{\epsilon}$ and $\mathbf{E}$ are no more subject to matrix inversion. 
The bias and the covariance of the estimate $\widehat{\bm{\theta}}$ are obtained from the definitions (\ref{eqn:biasdef}), and (\ref{eqn:covdef}). %The bias and the covariance of the estimate $\widehat{\bm{\theta}}$ are obtained from \begin{equation}  \mathbf{b} \left(\widehat{\bm{\theta}} \right) = \mathbf{\mu} - \widebar{\bm{\theta}}, \label{eqn:biasdef} \end{equation}  \begin{equation} \begin{aligned} \mathrm{\mathbf{Cov}} \left( \widehat{\bm{\theta}} \right) & = \mathbb{E} \left\{ \left( \widehat{\bm{\theta}} - \mathbf{\mu} \right)  \left( \widehat{\bm{\theta}} - \mathbf{\mu} \right)^\top \right\} . \end{aligned} \label{eqn:covdef} \end{equation} where $\mathbf{\mu} = \mathbb{E} \left\{ \widehat{\bm{\theta}} \right\}$ is the expected value of the estimate and $\widebar{\bm{\theta}}$ is the true value. 
Considering the structure of the EIV problem, the bias and the covariance of the estimate approximation (\ref{eqn:xhatexpW}) can be expressed as
\begin{equation} \begin{aligned} \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{C}^{-1} \left(  \left( \mathbf{K}^\top \bm{\Omega} \mathbf{B}_1 - \mathbf{B}_2 \right) \bm{\theta} - \left( \mathbf{K}^\top \bm{\Omega} \mathbf{B}_3 - \mathbf{B}_4 \right) \right), \end{aligned} \label{eqn:biasEW} \end{equation}
\begin{equation} \begin{aligned} \mathbf{C}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \mathbf{K}^\dagger \bm{\Omega} \left( \sigma_{\epsilon}^2 \mathbf{I}_{\color{blue}N\color{black}-n} + \mathbf{C}_1 - \mathbf{C}_2 - \mathbf{C}_2^\top \right) \bm{\Omega} \mathbf{K}^{\dagger \top} - \mathbf{b}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \mathbf{b}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right), \end{aligned} \label{eqn:varEW} \end{equation}
where $\mathbf{B}_1 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\Omega} \mathbf{E} \Big\}$, $\mathbf{B}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Omega} \mathbf{P}_\perp \mathbf{E} \Big\}$, $\mathbf{B}_3 = \mathbb{E} \Big\{ \mathbf{E} \mathbf{K}^\dagger \bm{\Omega} \bm{\epsilon} \Big\}$, $\mathbf{B}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Omega} \mathbf{P}_\perp \bm{\epsilon} \Big\}$, $\mathbf{C}_1 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\theta}^\top \mathbf{E}^\top \Big\}$, $\mathbf{C}_2 = \mathbb{E} \Big\{ \mathbf{E} \bm{\theta} \bm{\epsilon}^\top \Big\}$, $\mathbf{P}_\perp = \mathbf{I} - \mathbf{K} \mathbf{K}^\dagger \bm{\Omega}$, and $\mathbf{K}^\dagger$ is the pseudo-inverse matrix of $\mathbf{K}$.

The bias and covariance given by expressions (\ref{eqn:biasEW}) and (\ref{eqn:varEW}) depend on the unobservable true values $\bm{\theta}$ and $\mathbf{K}$.
The measured observations are in the sensor step response $\widetilde{\mathbf{y}}$, and from its observations we construct $\widetilde{\mathbf{K}}$ and compute $\widehat{\bm{\theta}}$.
The substitution of the measured data in the expressions gives an approximation of the estimation bias and covariance. 
We have then
\begin{equation} \begin{aligned} \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{C}}^{-1} \left(  \left( \widetilde{\mathbf{K}}^\top \bm{\Omega} \widetilde{\mathbf{B}}_1 - \widetilde{\mathbf{B}}_2 \right) \widehat{\bm{\theta}} - \left( \widetilde{\mathbf{K}}^\top \bm{\Omega} \widetilde{\mathbf{B}}_3 - \widetilde{\mathbf{B}}_4 \right) \right), \end{aligned} \label{eqn:biasSTW} \end{equation}
  \begin{equation} \begin{aligned} \widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) & \approx \widetilde{\mathbf{K}}^\dagger \bm{\Omega} \left( \sigma_{\epsilon}^2 \mathbf{I}_{\color{blue}N\color{black}-n} + \widetilde{\mathbf{C}}_1 - \widetilde{\mathbf{C}}_2 - \widetilde{\mathbf{C}}_2^\top \right) \bm{\Omega} \widetilde{\mathbf{K}}^{\dagger \top} - \widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right) \widetilde{\mathbf{b}}_{\mathrm{p}}^\top \left( \widehat{\bm{\theta}} \right), \end{aligned} \label{eqn:varSTW} \end{equation}
  where $\widetilde{\mathbf{B}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\Omega} \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_2 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Omega} \widetilde{\mathbf{P}}_\perp \mathbf{E} \Big\}$, $\widetilde{\mathbf{B}}_3 = \mathbb{E} \Big\{ \mathbf{E} \widetilde{\mathbf{K}}^\dagger \bm{\Omega} \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{B}}_4 = \mathbb{E} \Big\{ \mathbf{E}^\top \bm{\Omega} \widetilde{\mathbf{P}}_\perp \bm{\epsilon} \Big\}$, $\widetilde{\mathbf{C}}_1 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \widehat{\bm{\theta}}^\top \mathbf{E}^\top \Big\}$, $\widetilde{\mathbf{C}}_2 = \mathbb{E} \Big\{ \mathbf{E} \widehat{\bm{\theta}} \bm{\epsilon}^\top \Big\}$, and $\widetilde{\mathbf{P}}_\perp = \mathbf{I} - \widetilde{\mathbf{K}} \widetilde{\mathbf{K}}^\dagger \bm{\Omega}$.

The results of the expected values $\mathbf{B}_1$, $\mathbf{B}_2$, $\mathbf{B}_3$, $\mathbf{B}_4$, $\mathbf{C}_1$, and $\mathbf{C}_2$ can be found using Lemma \ref{lem:lemma1}, on page \pageref{lem:lemma1}.
The bias and covariance were obtained to extend the previous analysis conducted on EIV estimation problems without an imposed structure in \citet{Vaccaro94} and \citet{Stewart90SPT}.
It was shown that the bias and variance expressions (\ref{eqn:biasSTW}) and (\ref{eqn:varSTW}) are valid predictions of the first and second moments of the LS estimate of a Hankel structured EIV problem.
The problem formulated by the step input estimation method belongs to this type of structured EIV problems and we can use the derived expressions to find the bias and variance of the input estimate $\widehat{u}$.
The bias of the estimate $\widehat{\mathbf{u}}$ is the fist element of $\widetilde{\mathbf{b}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right)$ and the variance of $\widehat{\mathbf{u}}$ is the first element in the main diagonal of $\widetilde{\mathbf{C}}_{\mathrm{p}} \left( \widehat{\bm{\theta}} \right)$.

\subsection{Maximum-likelihood method}

% This means that the ML method implementation depends on the problem under study and not always can be estimated all the sensor model parameters.  
% The identifiability conditions of each sensor model parameter is the requirement to fulfill in order to be estimated with the ML method. %The affine input parameters $a$ and $b$ are the slope and the intercept of the straight line model $u(t) = at + b$.
